On semidefinite-representable sets over valued fields

### 2025-12-09 issac2026/main.tex
commit by Simone

\documentclass[a4paper,oneside,10pt]{article}

%%%\newenvironment{proof}{\hbox{}\vspace{-0.8cm} {\bf Proof:}}{\hfill $\Box$ \vspace{0.2cm}}

@@

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
%\newtheorem{algorithm}[theorem]{Algorithm}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{condition}[theorem]{Condition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{problem}[theorem]{Problem}
\newtheorem{example}[theorem]{Example}
\newtheorem{notation}[theorem]{Notation}
\usepackage{booktabs}

\usepackage{geometry}
\geometry{
 a4paper,
 %total={170mm,257mm},
 left=33mm,
 right=33mm,
 top=35mm,
 bottom=35mm
}

%\textheight235mm
%\textwidth165mm
%\voffset-10mm
%\hoffset-12.5mm
\parindent0.8cm
\parskip0mm

@@

%\usepackage{mathptmx}

% updates

@@

\usepackage{biblatex}
\addbibresource{../rapport/bibstage.bib}

\title{\Large \bf On semidefinite-representable sets over valued fields}

\begin{document}

\author{Corentin Cornou$^{1}$, Simone Naldi$^{2}$ and Tristan Vaccon$^{2}$}

\footnotetext[1]{ENS Paris-Saclay, Université Paris-Saclay, France.}
\footnotetext[2]{Université de Limoges, CNRS, XLIM, UMR 7252, F-87000 Limoges, France.}
%\footnotetext[3]{Sorbonne Université, CNRS, LIP6, Equipe PolSys, F-75005 Paris, France.}

\date{Draft of \today}

\maketitle

\begin{abstract}
  \noindent
  Real polyhedra and real spectrahedra (or their projections) are respectively the feasible sets
  of linear and semidefinite programming, and form the family of
  semidefinite-representable sets. This paper investigates versions of these sets and of
  the related optimization problems with data over a valued field $K$.
  For $K$-polyhedra and linear programming over $K$ we give an algorithm based on computation
  of Smith normal forms.
  We prove that basic properties of semidefinite-representable sets generalize to this setting.
  In particular, we showcase examples of non-polyhedral $K$-spectrahedra and of sets that are
  semidefinite representable over $K$ but are not $K$-spectrahedra.
\end{abstract}

%\tableofcontents

\section{Introduction}

The study of polyhedra and of linear programming (LP) is at the heart of convex geometry and optimization, inasmuch as several optimization problems can be cast as, or relaxed to LP. An important generalization of LP is semidefinite programming (SDP), namely the problem of minimizing a linear function over the set of positive semidefinite affine combinations of some real symmetric matrices. The latter problem has interested the community of real algebraic geometry and polynomial optimization, mostly because of the relaxation method known as moment-SOS (or Lasserre) hierarchy \cite{henrion2020moment}.

The main motivation for studying the effective aspects of SDP is the fact that, despite several properties of LP extend to SDP, the complexity status of SDP is, contrary to the LP case, widely open. Ramana's construction \cite{ramana1997exact} implies a series of general results: in the Blum-Shub-Smale real numbers machines model \cite{blum1989theory}, the feasibility of SDP is in $\text{NP} \cap \text{co-NP}$, whereas it is not known to be in NP in the Turing model; checking boundedness, attainability and optimality can be polynomially reduced to the feasibility of SDP \cite{ramana1997exact}. Assuming the input data belongs to an effective field $K$ (\emph{e.g.} $K = \QQ$) then a solution can be described and computed as element of a vector space over some algebraic extension of $K$. For exact computation, complexity bounds are known based either on general quantifier elimination \cite{ramana1997exact,porkolab1997complexity} and explicit algorithms exist based on a dedicated application of the so-called critical point method, and exploiting the determinantal structure of the constraints \cite{henrion2016exact,NALDI2018206}. A complexity analysis for the computation of an approximate solution of an SDP can be made based on the ellipsoid algorithm \cite{khachiyan1979polynomial,grotschel1981ellipsoid} or on interior-point methods \cite{de2016turing}; in this approximation framework, SDP is solvable in polynomial time in the input size (\emph{i.e.} matrix size, number of variables, bit size of the coefficients), $\log(\frac{1}{\varepsilon})$ (where $\varepsilon$ is the target precision) and $\log(\frac{R}{r})$, where $r,R$ are radii of balls respectively included in and containing the feasible set, cf. \cite{grotschel1981ellipsoid} or \cite[Thm.1.1]{de2016turing} (see also \cite[Sec.1.9]{deKlerk}).

In this paper we consider the situation where input data are defined over fields $K$ with a prescribed valuation map $\val$ on $K$. In this context a polyhedron is defined as a set of vectors such that some linear forms have nonnegative or infinite valuation ({\it cf.} \Cref{def_polyhedra}). The prototype of valued field we target is the field of $p$-adic numbers with $p$-adic valuation.
%%%\todo{Give motivations for p-adic precision of LP/SDP.}
Similar generalizations have been recently proposed in the context of tropical geometry \cite{iriarte2022polyhedral} and tropical methods applied to non-archimedean semidefinite programming \cite{allamigeon2016solving,allamigeon_tropical_2020}.

%\subsection{Context and motivations}
%Il faudra entre autre mentionner (et peut être motiver l'approche p-adique) :
%\begin{itemize}
%\item les questions ouvertes de complexité concernant la programmation semidefinie (pas clair si NP $\cap$ co-NP
%  dans le modele de Turing)
%\item les questions ouvertes géométriques (Conjecture de Lax)
%\end{itemize}
%\subsection{Main results}

%\newpage
\section{Preliminaries}

Throughout the paper, $(K,\val)$ refers to a complete valued field, namely $K$ is a field
and $\val : K \twoheadrightarrow \Gamma \cup \{+\infty\}$ is a valuation on $K$, $\Gamma$ an
ordered abelian group. The valuation is often assumed to be discrete ({\it i.e.} $\Gamma=\ZZ$)
with $\pi \in K$ a uniformizer. The valuation ring of $K$ is
$$
\OK = \{x \in K : \val(x) \geq 0\},
$$
$\kappa = \OK/\mathfrak{m}_K$ its residue field, with
$\mathfrak{m}_K = \{x \in K : \val(x) > 0\}$. An element of $\OK$ is called integral (at $\val$).
For $k \in \N$, we write $O(\pi^k)$ for $\pi^k \OK$.
A typical example of $K$ as above is the field of $p$-adic numbers 
$\QQ_p$ equipped with the $p$-adic valuation. For this example, we 
have $\OK = \ZZ_p$. Another example is the field of Laurent series
$K=\QQ(\!(t)\!)$ equiped with $t$-adic valuation,
in which case $\OK = \QQ \llbracket t \rrbracket$.
We refer to \cite{Serre:1979,engler2005valued} for a
general introduction to such fields
and to \cite{caruso_computations_2017}
for an introduction to effective computations over $p$-adic numbers.

The $K$-vector space of polynomials of degree $\leq d$ on $x=(x_1,\ldots,x_n)$
with coefficients in $K$ is denoted by $K[x]_{d}$. The ring of matrices of
size $p \times q$ with entries in a ring $R$ is $R^{p \times q}$ and the general
linear group in $R^{n \times n}$ is $\GL_n(R)$.
For the sake of ease of writing, block-diagonal matrices with blocks
$B_1,\ldots,B_d$ are denoted by $\diag(B_1, \ldots, B_d)$, the null matrix of
size $p \times q$ by ${\bf 0}_{p,q}$, and the identity matrix of size $t \times t$ by $I_t$.

We recall the definition of Smith Normal Form, in our context.
\begin{definition}[Smith Normal Form]\label{smith_nf}
  Assume $(K,\val)$ is a discrete valued field with uniformizer $\pi$.
  Let $M \in K^{m \times n}$ be of rank $r$. There exist unique integers
  $a_1 \leq \cdots \leq a_r$, and
  matrices $P \in \GL_n(\OK)$ and $Q \in \GL_m(\OK)$, such that
  $$
  \mathrm{SNF}(M) := QMP^{-1} = \diag(\pi^{a_1}, \pi^{a_2}, \ldots, \pi^{a_r}, {\bf 0}_{m-r,n-r}).
  $$
  The matrix $\mathrm{SNF}(M)$ is called the \emph{Smith Normal Form (SNF)} of $M$, and the elements $\pi^{a_i}$ the
  \emph{invariant factors}.
\end{definition}
\begin{remark}
  The valuation of the first invariant factor of $\mathrm{SNF}(M)$ of a matrix
  $M \in K^{n \times n}$ is the minimum of the valuations of the entries of
  $M$. In particular, $\mathrm{SNF}(M) \in \OK^{n \times n}$.
\end{remark}

Let us now give a short introduction to classical ({\it i.e.} real) polyhedra and spectrahedra.
Let $\sym_d(\R) \subset \R^{d\times d}$ be the vector space of $d \times d$ real symmetric
matrices. A matrix $M \in \sym_d(\R)$ is \emph{positive semidefinite} ($M \succeq 0$) whenever the quadratic form $x \mapsto x^TMx$ is nonnegative for all $x\in \R^n$. By the Spectral Theorem, $M \succeq 0$ if and only if the eigenvalues of $M$ are nonnegative, and this is equivalent to all the principal minors of $M$ being nonnegative. The set $\sym_d^+(\R) \subset \sym_d(\R)$ of positive semidefinite matrices is a closed convex cone with non-empty interior in $\sym_d(\R)$.

Let $A_0,A_1,\ldots,A_n \in \sym_d(\R)$, and let $\calL = A_0+\span{A_1,\ldots,A_n}$ be the affine space
containing $A_0$ and with direction the vector space spanned by $A_1,\ldots,A_n$ (which we might assume
to be linearly independent). The intersection
$$
\calL \cap \sym_d^+(\R) = \left\{A(x) := A_0+\sum_i x_i A_i \in \sym_d(\R) :
A(x) \succeq 0\right\},
$$
or equivalently, its pre-image $S = \{x \in \R^n : A(x) \succeq 0\}$ under the map $x \mapsto A(x)$, is called
a \emph{real spectrahedron}. As (linear preimage of) affine sections of $\sym_d^+(\R)$, real spectrahedra
are closed convex sets, but might have empty interior in $\calL$ (resp. in $\R^n$).
Moreover, they are basic semialgebraic sets, indeed they are defined by sign conditions on the principal
minors of the defining matrix as well as sign conditions on the coefficients of its characteristic polynomial
(cf. \Cref{ell3}).

\begin{example}\label{ell3}
  Let $d=n=3$. The $3$-elliptope is the three-dimensional spectrahedron given by
  $$
  \mathcal{E}_3 :=
  \left\{
  x
  \in \R^3 :
  \begin{bmatrix}
    1 & x_1 & x_2 \\
    x_1 & 1 & x_3 \\
    x_2 & x_3 & 1
  \end{bmatrix}
  \succeq 0
  \right\}
  =
  \left\{
  x
  \in \R^3 :
  \begin{array}{r}
    3-x_1^2-x_2^2-x_3^2 \geq 0 \\
    1+2x_1x_2x_3-x_1^2-x_2^2-x_3^2 \geq 0
  \end{array}
  \right\}.
  $$
  The polynomials on the right are the (non-constant) coefficients of the univariate polynomial
  $p(t) = \det(A(x)+t I_3)$,
  whose roots are the opposites of the eigenvalues of the defining matrix. Equivalently $\mathcal{E}_3$
  is defined by positivity of the principal minors of the defining matrix:
  $$
  1-x_1^2 \geq 0, \,\,\, 1-x_2^2 \geq 0, \,\,\, 1-x_3^2 \geq 0 \,\,\, \text{ and } \,\,\, 1+2x_1x_2x_3-x_1^2-x_2^2-x_3^2 \geq 0.$$ \exend
\end{example}

The optimization problem of miminizing a linear function over a real spectrahedron is called \emph{semidefinite programming} (SDP).
%Given $c \in \R^n$ and $A_0,A_1,\ldots,A_n \in \sym_d(\R)$, the corresponding semidefinite program is defined (in standard dual form) as
%\begin{equation}
%  \label{SDP}
%\begin{array}{rcll}
%  p^* & := & \inf_{x \in \R^n} & \left\langle c, x \right\rangle \\
%  &    & \text{s.t.}         & A_0+\sum_{i=1}^n x_i A_i \succeq 0.
%\end{array}
%\end{equation}
{\it Real polyhedra} (subsets of $\R^n$ defined by finitely many affine inequalities) are special cases of real
spectrahedra, where the defining matrices $A_0, \ldots, A_n$ can be chosen diagonal, so that
$A = \diag(\ell_1,\ldots, \ell_d)$. In other words, SDP is a generalization of Linear Programming (LP).

% IN INTRODUCTION
%In a fixed-precision model of computation, SDP is solvable in polynomial time in the input size (matrix size,
%number of variables, bit size of the coefficients), in $\log(\frac{1}{\varepsilon})$ (where $\varepsilon$ is the
%precision) and $\log(\frac{1}{R})$, where $R$ is an {\it a priori} upper bound on the norm of a solution,
%cf. \cite[Sec.1.9]{deKlerk}. In exact arithmetic, the complexity status of SDP is essentially open but complexity
%bounds are known based either on general quantifier elimination \cite{ramana1997exact,porkolab1997complexity}
%or by exploiting the determinantal structure of the constraints \cite{henrion2016exact}.

A more general class of convex semialgebraic sets are obtained by linear projections of real spectrahedra
called \emph{semidefinite representable sets}: these are sets of the form
$$
\left\{x = \left[\begin{smallmatrix} x_1 \\ \vdots \\ x_n \end{smallmatrix}\right] \in \R^n : \exists\,y\in\R^m, \, A_0 + \sum_{i=1}^n x_i A_i + \sum_{j=1}^m y_j B_j \succeq 0\right\}.
$$
When $m>0$, the set is also called a \emph{spectrahedral shadow}.

\begin{example}
\label{fermat_quartic}
The basic closed semialgebraic set $\mathcal{R} = \{(x_1,x_2) \in \R^2 : 1-x_1^4-x_2^4 \geq 0\}$ is not a spectrahedron
but it is semidefinite representable:
$$
\mathcal{R} = \left\{\begin{bmatrix} x_1 \\ x_2 \end{bmatrix} \in \R^2 :
\exists\,
\begin{bmatrix} y_1 \\ y_2 \end{bmatrix} \in \R^2, \,
\diag\left(
\begin{bmatrix}
  1+y_1 & y_2 \\
  y_2 & 1-y_1
\end{bmatrix},
\begin{bmatrix}
  1 & x_1 \\
  x_1 & y_1
\end{bmatrix},
\begin{bmatrix}
  1 & x_2 \\
  x_2 & y_2
\end{bmatrix}
\right)
\succeq 0
\right\}
$$
\end{example}

\section{Polyhedra}

In this section, we define polyhedra over a complete discrete valued field $K$ with valuation $\val$ and a uniformizer $\pi$, and we prove that the class of polyhedra is closed under linear transformations, as in the real case. We end by presenting an algorithm for solving linear optimization problems over $K$.

\begin{definition}
  \label{def_polyhedra}
  A \emph{polyhedron} is a subset $\PP \subset K^n$ of the form
  $$
  \PP = \{x \in K^n : \val(\ell_i(x)) \geq 0, i=1,\ldots,d \text{ and }
  \val(m_j(x)) = +\infty, j=1,\ldots,e\},
  $$
  for some $\ell_1,\ldots,\ell_d,m_1,\ldots,m_e \in K[x]_1$.
\end{definition}

%%We recall that that the smallest affine space of $K^n$ containing a set $S \subset K^n$ is called its \emph{affine hull} and denoted $\aff{S}$. The assumption on forms $\ell_i$ in \Cref{def_polyhedra} implies that the constraint $\val(m_1 \oplus \cdots \oplus m_e) = +\infty$ defines exactly the affine hull of $P$, in other words $\aff{P} = \{x \in K^n : m_j(x) = 0, j=1,\ldots,e\}$ \simone{(il nous faudrait une preuve, du genre: $\val(m(x))=+\infty$ pour tout $x \in P$ si et seulement si $m$ est une combinaison affine de $m_1,\ldots,m_e$.)}
%%The \emph{dimension} of $P$ is defined as the dimension of $\aff{P}$ (for $e=0$, $\aff{P}=K^n$ and $P$ has dimension $n$). The \emph{relative interior} of $P$ is defined as $\relint{P} = \{x \in P : \val(\ell_1 \oplus \cdots \oplus \ell_d) < +\infty\}$ (\simone{pas clair si besoin de dimension et relint}).
Remark that \Cref{def_polyhedra} includes the case $d=0$ corresponding to affine spaces, as in the real case.
Moreover by definition the class of polyhedra is closed under intersection. Without loss of generality, one may
assume in \Cref{def_polyhedra} that for every $i=1,\ldots,d$ there exists $x \in P$ such that
$\val(\ell_i(x))<+\infty$.

\begin{example}
  The \emph{closed unit ball} $\OK^n=B_\infty(K^n)$ in $K^n$ for the supremum norm $||\cdot||_{\infty}$ is the
  polyhedron defined as the set of $x=(x_1,\ldots,x_n) \in K^n$ such that $||x||_{\infty} =
  \max_i\{\vert x_i \vert\} \leq 1$, in other words, such that $\val(x_i) \geq 0$, for all $i=1, \ldots, n$.
  As such, $\OK^n$ is the polyhedron in $K^n$ defined by the affine polynomials
  $\ell_i(x)=x_i$, $i=1,\ldots, n$, and $e=0$. %%%%%(its affine hull is indeed the whole $K^n$).
\end{example}

\subsection{Projections and direct image of polyhedra}

%{\color{blue} Normalement avec la nouvelle definition, ça devrait aller, il faut par contre
%  partir d'un polyedre defini aussi par des égalités. Peut être il faudrait aussi
%  diviser la proposition en plusieures parties pour la rendre plus lisible, par exemple une
%  première proposition avec la projection.}

In this section, we prove that the class of polyhedra is stable
by applying affine maps:

\begin{theorem}[DI] 
Let $f : K^n \to K^m$ be an affine map and $\PP$ a polyhedron in $K^n.$
Then $f(\PP)$ is a polyhedron. \label{theo:DI}
\end{theorem}

We prove the result by reducing to particular cases.
The strategy to deduce the main theorem DI 
is given in \Cref{fig:di-properties}.

\begin{figure}[!ht] 
\centering
\begin{tikzpicture}[
  node distance=1cm and 1cm,
  box/.style={draw, rounded corners, minimum width=2.2cm, minimum height=1cm, align=center},
  thickarrow/.style={->, >=Stealth, line width=1.2pt},
]
% Nodes
\node[box] (DIdiag) {DI${}_{diag}$};
\node[box, right=of DIdiag] (DIproj) {DI${}_{proj}$};
\node[box, right=of DIproj] (DIl) {DI${}_{l}$};
\node[box, right=of DIl] (DIi) {DI${}_{i}$};
\node[box, right=of DIi] (DI) {DI};
\node[box, above=of DIproj] (DIimm) {DI${}_{imm}$};
\node[box, below=of DIproj] (DIaut) {DI${}_{aut}$};
% Arrows
\draw[thickarrow] (DIdiag) -- (DIproj);
\draw[thickarrow] (DIproj) -- (DIl);
\draw[thickarrow] (DIl) -- (DIi);
\draw[thickarrow] (DIi) -- (DI);
\draw[thickarrow] (DIimm) -- (DIl);
\draw[thickarrow] (DIaut) -- (DIl);
\end{tikzpicture}
\caption{Strategy of proof of Theorem \ref{theo:DI}.}
\label{fig:di-properties}
\end{figure}

We begin with the case of linear maps in $\GL_n(O_K)$.

\begin{lemma}[DI${}_{aut}$]
Let $f \in GL_n(O_K)$ and $\PP$ a polyhedron in $K^n$.
Then $f(\PP)$ is a polyhedron. \label{lem:DIaut}
If in addition, $\PP$ is defined only by inequalities,
then so is $f(\PP)$.
\end{lemma}
\begin{proof}
If $\PP$ is defined (in matrix form) by \[\left\lbrace X \in K^n,  AX+B \geq 0, CX+D=0 \right\rbrace\] and $U$ is the matrix of $f$ in the canonical basis of $K^n$, 
let $\Qc$ be defined by \[\left\lbrace Y \in K^n,  AU^{-1} Y+B \geq 0, CU^{-1} Y+D=0 \right\rbrace.\]
Then it is clear that $\Qc$ is a polyhedron and $f(\PP)=\Qc.$
It is also clear that if $\PP$ is defined only by inequalities,
then so is $f(\PP)$.
\end{proof}

If $f$ is diagonal, we can also conclude.

\begin{lemma}[DI${}_{diag}$]
Let $f : K^n \to K^n$ be a linear map and $\PP$ a polyhedron in $K^n$ defined only by inequalities. We assume that the matrix of $f$ in the canonical basis of $K^n$ is diagonal.
Then $f(\PP)$ is a polyhedron. \label{lem:DIdiag}
\end{lemma}
\begin{proof}
As in the proof of Lemma \ref{lem:DIaut} DI${}_{aut}$, a trivial change
of variables is enough to conclude.
\end{proof}

We now treat the case where $f$ is a projection erasing the last coordinate.

\begin{lemma}[DI${}_{proj}$]
Let $f : K^n \to K^{n-1}, \: (x_1,\dots,x_{n-1},x_n) \mapsto (x_1,\dots,x_{n-1})$ and $\PP$ a polyhedron in $K^n$ defined only by inequalities. 
Then $f(\PP)$ is a polyhedron defined only by inequalities. \label{lem:DIproj}
\end{lemma}
\begin{proof}
We assume that $\PP \neq \emptyset$.
Let us write the conditions defining $\PP$
in matrix form:
$\val(A X+B) \geq 0$ 
with $A \in M_{q,n}(K)$, $B \in K^q$
and $X=(x_1,\dots,x_n)^\intercal$.

We can compute a Smith Normal Form
of the $q \times (n-1)$ matrix defined by the 
$n-1$ first columns of $A$
to obtain a decomposition of the form:
\[ Q_A A P_A^{-1} = \begin{bmatrix}
\delta_1	& 		& 			&   &		 &  &a_1	\\
			& \ddots& 			& 0	&		 &	&		\\
			&		& \delta_l  &   & 		 &	&		\\
			&		&			&0  & 		 &	& 		\\
			&		&			&   & \ddots &	&		\\
			&		&			&	&		 & 0&a_{n-1} \\
			&		&			&	&		 &	&a_n 	\\
			&		&			&	&		 &	&    	\\
			&		&			&	&		 &	&a_q 	\\			
\end{bmatrix},\]
with $Q_A \in \GL_q(\OK)$ and $P_A \in \GL_n(\OK)$ such that $P_A(e_n)=e_n.$

Multiplying on the left by $Q_A^{\pm 1}$ does not
modify whether the valuation inequalities are satisfied.
Moreover, thanks to Lemma \ref{lem:DIaut} DI${}_{aut}$,
$\PP$ can be written
$\PP=P_A^{-1}(\PP_2)$ for some polyhedron $\PP_2 \subset K^n$
and we are reduced to proving that
$f(\PP_2)$ is a polyhedron for $p$ the projection
$K^n \rightarrow K^{n-1}$ and $\PP_2$
defined by the following linear matrix inequality (for some $B_2 \in K^q$):
\[  \begin{bmatrix}
\delta_1	& 		& 			&   &		 &  &a_1	\\
			& \ddots& 			& 0	&		 &	&		\\
			&		& \delta_l  &   & 		 &	& \vdots\\
			&		&			&0  & 		 &	& 		\\
			&		&			&   & \ddots &	&		\\
			&		&			&	&		 & 0&a_{n-1} \\
			&		&			&	&		 &	&a_n 	\\
			&		&			&	&		 &	& \vdots   	\\
			&		&			&	&		 &	&a_q 	\\			
\end{bmatrix} \begin{bmatrix} x_1 \\ \\ \\ \\ \\ \vdots \\ \\ \\ \\ x_n \\ \end{bmatrix} + B_2 \geq 0.\]

We remark that 
$a_s x_n + b_s \geq 0$
if and only if $x_n = -a_s^{-1} b_s + O(\pi^{- \val(a_s)}).$
This exactly corresponds to $x_n \in B(-a_s^{-1} b_s,\val(a_s)).$
Since the conditions have to be compatible (we have assumed
$\PP$ not empty), and thanks to ultrametricity,  
the intersection of these conditions for 
$s \in \llbracket l+1,q \rrbracket$
is then given by the smallest of these balls.
Let us assume it is the one defined for $l=n.$
The only condition on $x_n$
is then $\val(a_nx_n +b_n) \geq 0.$

Thanks to Lemma \ref{lem:DIdiag} DI${}_{diag}$,
we can further assume that the $\delta_i$'s and $a_n$ are all $1$'s
and are reduce to $\PP_2$ defined by the linear matrix inequality:
\[  \begin{bmatrix}
1	& 		& 			&   &		 &  &a_1	\\
			& \ddots& 			& 0	&		 &	&		\\
			&		& 1  &   & 		 &	& \\
			&		&			&0  & 		 &	& 	\vdots	\\
			&		&			&   & \ddots &	&		\\
			&		&			&	&		 & 0&a_{n-1} \\
			&		&			&	&		 &	&1 	\\

\end{bmatrix} \begin{bmatrix} x_1 \\ \\  \\ \\ \vdots \\  \\ \\ x_n \\ \end{bmatrix} + \begin{bmatrix} b_1 \\ \\ \\  \\ \vdots  \\ \\ \\ b_n \\ \end{bmatrix}  \geq 0.\]
Finally, we can also assume that, up to a permutation of the variables, $\val(a_1)= \min_{i=1}^{n-1} \val(a_i).$
We distinguish two cases: $\val(a_1) \geq 0$ and $\val(a_1)<0.$

Let us first assume that $\val(a_1) \geq 0$.

Let $x \in \PP_2$.
Then $x_n=-b_n+O(1)$
and $x_1+a_1 x_n+b_1=O(1).$
Consequently, $x_1-a_1 b_n +b_1=O(1)+O(\pi^{\val(a_1)})=O(1).$
Similarly, for all $i \in \llbracket 2,n-1 \rrbracket,$
$x_i-a_ib_n+b_i=O(1)+O(\pi^{\val(a_i)})=O(1).$

We define the polyhedron $\Qc \subset K^{n-1}$ by the following inequalities:
 for all $i \in \llbracket 1,n-1 \rrbracket,$
$\val(x_i-a_ib_n+b_i) \geq 0$.

Then the previous computations prove that $p(\PP_2)\subset \Qc.$
Conversely, let $(x_1,\dots,x_{n-1}) \in \Qc.$
Then let $x_n := -b_n.$

We check that for $i \in \llbracket 1,n-1 \rrbracket,$
$x_i+a_ix_n+b_i=x_i-a_ib_n+b_i=O(1),$
and $x_n+b_n=O(1)$. Hence $(x_1,\dots,x_n) \in \PP_2$
which means that $(x_1,\dots,x_{n-1}) \in p(\PP_2).$
Thus $p(\PP_2)= \Qc$ and $p(\PP_2)$ is a polyhedron.

We now deal with the second case, with $\val(a_1)<0.$
Let $x \in \PP_2$.
Then $x_n=-b_n+O(1)$
and $x_1+a_1 x_n+b_1=O(1).$
Consequently, $x_1-a_1 b_n +b_1=O(1)+O(\pi^{\val(a_1)})=O(\pi^{\val(a_1)}).$
Thus $a_1^{-1}x_1- b_n +a_1^{-1}b_1=O(1)$.
%and $x_n=-a_1^{-1}(x_1+b_1)+O(1).$
In addition, $x_n=-a_1^{-1}(x_1+b_1)+O(\pi^{-\val(a_1)}).$
We plug this equality in 
$x_i+a_i x_n+b_i=O(1)$ to get
$x_i-a_1^{-1}a_i(x_1+b_1)+b_i=O(1)+O(\pi^{\val(a_i)-\val(a_1)})=O(1)$
using that $\val(a_i) \geq \val (a_1).$

We define the polyhedron $\Qc \subset K^{n-1}$ by the following inequalities:
$\val(a_1^{-1}x_1- b_n +a_1^{-1}b_1) \geq 0$ and for all $i \in \llbracket 2,n-1 \rrbracket,$
$\val(x_i-a_1^{-1}a_i(x_1+b_1)+b_i) \geq 0.$

Then the previous computations prove that $p(\PP_2)\subset \Qc.$
Conversely, let $(x_1,\dots,x_{n-1}) \in \Qc.$
Let $x_n := -a_1^{-1}(x_1+b_1).$
Then
$x_1+a_1 x_n+b_1=0=O(1).$
Let $i \in \llbracket 2,n-1 \rrbracket,$
then $x_i+a_i x_n+b_i=x_i-a_1^{-1}a_i(x_1+b_1)+b_i=O(1)$
since $(x_1,\dots,x_{n-1}) \in \Qc.$
Finally, 
$x_n+b_n=-a_1^{-1}(x_1+b_1)+b_n=O(1)$ thanks to the inequality satisfied by $x_1.$
Hence $(x_1,\dots,x_n) \in \PP_2$
which means that $(x_1,\dots,x_{n-1}) \in p(\PP_2).$
Thus $p(\PP_2)= \Qc$ and $p(\PP_2)$ is a polyhedron,
and it has been proved in both cases.

Therefore, $p(\PP)$ is a polyhedron (and inequalities are enough to define $p(\PP_2)$ as a polyhedron).
\end{proof}

\begin{lemma}[DI${}_{imm}$]
  Let $f : K^l \to K^n$ be the immersion $(x_1,\dots, x_l) \mapsto (x_1,\dots, x_l,0,\dots, 0)$ and $\PP$
  a polyhedron in $K^l$. Then $f(\PP)$ is a polyhedron. \label{lem:DIimm}
\end{lemma}
\begin{proof}
Adding the equalities $x_{l+1}=0, \dots, x_n=0$ to the equations and inequations in
$x_1,\dots, x_l$ defining $\PP$, the result is clear.
\end{proof}

We can now generalize to linear mappings and
polyhedron defined only by inequalities.

\begin{proposition}[DI${}_l$]
Let $f : K^n \to K^m$ be a linear map and $\PP$ a polyhedron in $K^n$ defined only by inequalities.
Then $f(\PP)$ is a polyhedron. \label{prop:DIl}
\end{proposition}
\begin{proof}
We assume that $\PP \neq \emptyset$, and we identify $f$ with its matrix in $\allmat_{n,m}(K)$.
We assume that $\PP$ is defined as $\left\lbrace x \in K^n : \val (l_i (x) ) \geq 0, \textrm{ for } i \in I\right\rbrace$ for some
finite sets $I \subset \N$ and $l_i \in K[x]_1$, for $i \in I$ .

Thanks to Definition \ref{smith_nf}, (the matrix of) $f$
can be written $f=Q^{-1} \Delta P$
with $P \in \GL_n(\OK),$ $Q \in \GL_m(\OK)$
and $\Delta = \mathrm{SNF}(f)\in \allmat_{n,m}(K)$ of the form $\diag(\lambda_1, \ldots, \lambda_r, {\bf 0}_{\min(m,n)-r})$
  where ${\bf 0}_{\min(m,n)-r}$ is a null square matrix of size $\min(m,n)-r$ for $\lambda_1,\dots,\lambda_r$ non-zero and $r \in \N.$

By Lemma \ref{lem:DIaut} DI${}_{aut}$,
$P(\PP)$ is a polyhedron defined only by inequalities.

Let $\Qc= P(\PP)$
We now prove that $\Delta (\Qc)$ is a polyhedron.
One can write that $\Delta$ is the composition
of the immersion 
$K^r \rightarrow K^n$
$(x_1,\dots,x_r)\mapsto (x_1,\dots,x_r,0,\dots,0),$ 
 the diagonal invertible mapping
$K^l \rightarrow K^l$,
$(x_1,\dots,x_r)\mapsto (\delta_1 x_1,\dots,\delta_r x_r)$,
and the projections
$(x_1,\dots,x_{r+1})\mapsto (x_1,\dots,x_{r}),$
 $(x_1,\dots,x_{r+2})\mapsto (x_1,\dots,x_{r+1}),$
 $\dots,$
 $(x_1,\dots,x_{n})\mapsto (x_1,\dots,x_{n-1}).$
 Thus, thanks to Lemmas \ref{lem:DIimm} DI${}_{imm}$, \ref{lem:DIproj} DI${}_{proj}$ and \ref{lem:DIaut} DI${}_{aut}$, $\Delta (\Qc)$ is a polyhedron.
 One can note that it is only when applying the immersion as the last mapping in the composition chain that equalities are needed to define the last polyhedron.

Finally, thanks to Lemma \ref{lem:DIaut} DI${}_{aut}$,
$Q^{-1} \left( \Delta (\Qc) \right)$ is a polyhedron.
This concludes the proof.
\end{proof}

We can generalize from linear mappings to affine mappings.
\begin{proposition}[DI${}_i$]
Let $f : K^n \to K^m$ be an affine map and $\PP$ a polyhedron in $K^n$ defined only by inequalities.
Then $f(\PP)$ is a polyhedron. \label{prop:DIi}
\end{proposition}
\begin{proof}
Let $P \in K^m$. One can see by immediate change of variables that $\Qc$ is a polyhedron in $K^m$ if and only if
$\Qc+P$ is a polyhedron in $K^m$.
Consequently, translating by $-f(O)$ in the codomain, one can reduce
to the case of Proposition \ref{prop:DIl}   DI${}_l$.
\end{proof}

We can now conclude with the proof of the main result
of this section.

>>>>

\documentclass[sigconf]{acmart}

@@

@@

@@

%\usepackage{biblatex}
%\addbibresource{../rapport/bibstage.bib}

\title{On semidefinite-representable sets over valued fields}

\author{Simone Naldi}
\affiliation{
    \institution{Université de Limoges, CNRS, XLIM}
    \city{Limoges}
    \country{France}
}

\author{Tristan Vaccon}
\affiliation{
    \institution{Université de Limoges, CNRS, XLIM}
    \city{Limoges}
    \country{France}
}

\author{Corentin Cornou}
\affiliation{
    \institution{}
    \city{}
    \country{France}
}

\copyrightyear{2026}
\acmYear{2026}
\setcopyright{cc}
\setcctype{by}
\acmConference[ISSAC '26]{International Symposium on Symbolic and Algebraic Computation}{Date}{Place}
\acmBooktitle{International Symposium on Symbolic and Algebraic Computation (ISSAC '26), date, place}\acmDOI{}
\acmISBN{}

%\begin{CCSXML}
%<ccs2012>
%   <concept>
%       <concept_id>10010147.10010148.10010149.10010150</concept_id>
%       <concept_desc>Computing methodologies~Algebraic algorithms</concept_desc>
%       <concept_significance>500</concept_significance>
%       </concept>
% </ccs2012>
%\end{CCSXML}

%\ccsdesc[500]{Computing methodologies~Algebraic algorithms}

\keywords{keywords}

\begin{abstract}
  Real polyhedra and real spectrahedra (or their projections) are respectively the feasible sets
  of linear and semidefinite programming, and form the family of
  semidefinite-representable sets. This paper investigates versions of these sets and of
  the related optimization problems with data over a valued field $K$.
  For $K$-polyhedra and linear programming over $K$ we give an algorithm based on computation
  of Smith normal forms.
  We prove that basic properties of semidefinite-representable sets generalize to this setting.
  In particular, we showcase examples of non-polyhedral $K$-spectrahedra and of sets that are
  semidefinite representable over $K$ but are not $K$-spectrahedra.
\end{abstract}

\begin{document}
\maketitle

\bibliographystyle{ACM-Reference-Format}
\bibliography{refs}
